\svnInfo $Id$

\section{Addressing Design}
\label{sec:addressing}

One of the \bus design points is to serve as a system interconnect for a
physically constrained system. With that in mind, \bus attempts to optimize
for a system of complex connected components. In particular, \bus expects that
an individual member node may itself be composed of multiple functional units,
which may each be individually addressed.

Supporting multiple, addressable functional units with only one set of exposed
\bus signal pins could be done by instantiating several copies of a
member node on each die, but this requires a bus interface module for every
functional unit within an integrated chip. Instead, the bottom four bits of
addresses are reserved to identify functional units. Nodes with more than
sixteen functional units will require multiple addresses.

\subsection{Address Types}
\label{sec:addressing-types}

\bus defines the term {\em prefix} to refer to the portion of the address that
specifies which node is being addressed and reserves the term {\em address} to
refer to a complete address---one that specifies a functional unit within a
node.

Every \bus node has two prefixes, a {\em full} prefix and a {\em short} prefix.
A short prefix is 4~bits long and a full prefix is 20~bits long.
A short address is the composition of short prefix and a functional
unit identifier. A full address is the composition of a header, a full prefix,
and a functional unit identifier.
To distinguish full and short addresses, the short prefix {\tt 0b1111} is
reserved to identify a full address. The first four bits of the full address
header is thus always {\tt 0b1111}.


\subsection{Full Prefix Assignment}
\label{sec:addressing-full}
~

\begin{bytefield}[bitwidth=1.4em]{32}
  \bitheader{0-31} \\
  \colorbitbox{lightgreen}{4}{0xF}
  \colorbitbox{lightergreen}{4}{0x0 (RSVD)}
  \bitbox{20}{Full Prefix}
  \colorbitbox{lightcyan}{4}{Func Unit ID} \\
  \bitbox[t]{8}{$\underbrace{\hspace{13em}}_{\text{\normalsize Header}}$}
\end{bytefield}

The purpose of full prefixes is to serve as a unique identifier for a node,
akin to a product identifier. Full prefixes do not distinguish instantiations
of a node, that is, multiple copies of a unique part will all have the same
full prefix. This implies that multiple nodes in a single \bus instantiation
may have the same full prefix. Bits~24-27 of the full address are reserved for
other purposes
(\ref{sec:todo-extensions-resume}~\nameref{sec:todo-extensions-resume}). The
remaining range, bits~4-23, are available to be utilized as full prefixes.
If a node has more than 16 functional units, it may have multiple full
prefixes. These full prefixes should be sequential.

\begin{quote}
\begin{itemize}
\item The full prefix {\tt 0x00000} is reserved as the broadcast prefix.
\end{itemize}
\end{quote}

\noindent
The full prefixes ranging {\tt 0x00000-0x0000F} are reserved for legacy M3
devices. \hl{An allocation scheme for new full prefixes is currently
undefined.}

\subsection{Short Prefix Assignment}
\label{sec:addressing-short}
~

\begin{bytefield}[bitwidth=1.5em]{8}
  \bitheader{0-7} \\
  \bitbox{4}{Short Prefix}
  \colorbitbox{lightcyan}{4}{Func Unit ID}
\end{bytefield}
\medskip

The purpose of short prefixes is to uniquely identify nodes in an \bus system.
Multiple nodes in a \bus instantiation {\bf must not} have the same short
prefix.

\begin{quote}
\begin{itemize}
\item The short prefix {\tt 0b0000} is reserved as the broadcast prefix.
\item The short prefix {\tt 0b1111} is reserved to distinguish full addresses.
\end{itemize}
\end{quote}

\noindent
This leaves a remainder of 14 unique short prefixes. These short prefixes map
to actual nodes instantiated in a \bus system. If there are multiple copies of
the same node type (e.g. several external memory chips), each instance is
given a unique short prefix. If a node has greater than 16 functional units
it will be assigned multiple short prefixes, each mapping to one of the node's
full prefixes.

In a \bus instantiation, short prefixes are assigned dynamically. Out of
reset, all member nodes are assigned the short prefix
\nameref{sec:spec-unassigned-short-prefix}. After system power-on, some node
shall send a series of \nameref{cmd:enumerate-node} commands. \bus does not
require that the mediator node send the \nameref{cmd:enumerate-node} commands,
however any node containing the mediator node should be capable of enumerating
bus members. Any node capable of performing enumeration {\bf must} be capable
of snooping another node's performing of enumeration and {\bf must} update its
local enumeration state from the snooped data.

Short prefixes should be assigned in ascending numerical order beginning with
prefix {\tt 0b0010}.
It is the responsibility of the
node(s) performing enumeration to ensure no duplicate short prefixes are
assigned.
As enumeration assignments are resolved by the
topological priority arbitration, the short prefix assigned to a node should
also correspond to its topological priority.
This constraint may be violated in systems that do not enumerate prefixes in
order, that use the
\nameref{cmd:invalidate-prefix} command to re-order short prefixes, or
that use a \nameref{sec:addressing-static-short-prefix} where the statically
assigned prefixes do not match the topology. Short prefixes cannot be used to
affect priority. Priority is determined exclusively by physical topology.

Short prefix assignments {\bf must} be preserved when nodes are powered gated.

\subsubsection{Static Short Prefix Assignment}
\label{sec:addressing-static-short-prefix}
All \bus nodes {\bf must} support dynamic short prefix assignments. A node may
have a default short prefix assignment. Upon the receipt of any
\nameref{cmd:enumerate-node} commands, a node with a default short prefix {\bf
must} immediately invalidate its default short prefix and participate in the
enumeration process. A node with a default short prefix that receives an
\nameref{cmd:invalidate-prefix} command shall set its short prefix to
\nameref{sec:spec-unassigned-short-prefix}. A node with a default short prefix
shall otherwise make no distinction from a dynamically assigned short prefix
(e.g., it shall respond to a \nameref{cmd:query-devices} command with its
current short prefix, the static short prefix).
\hl{An allocation scheme for default short prefixes is not yet defined.}

