\section{Test Cases}
\label{appx-tests}
This section documents some of the trickier corner cases in \bus design. This
is by no means a comprehensive list of tests, but each of these cases should
be tested to ensure compatibility across \bus implementations.

For most tests, we recommend that nodes are assigned a
\nameref{sec:addressing-static-short-prefix} so that enumeration can be
skipped.

\subsection{Registers}

\subsubsection{Dump Register Content to Memory and Restore to Other Registers}
\label{test:reg-dump-restore}

\begin{quote}
  \texttt{A) 4 Nodes: N1 -> N2 -> N3 -> MED (-> N1)}
\end{quote}

This test dumps a node's registers to the memory of another node by requesting
a register read that issues a streaming memory write. Some of these dumped
registers are then written to different registers on another node by using a
streaming memory read that issues a register write.

\begin{enumerate}
  \item Begin with N1's register data full of known, non-zero values. N2 and
    N3 registers and memory should be zeroed.
  \item N2 configures the memory stream address for N3:
    \subitem \nameref{cmd:mem-stream-conf}: \hl{0x3X + 0xXXXXXXXX}
  \item N2 requests that N1 copy 19 registers to N3's memory. Choose register
    addresses such that both read and destination addresses wrap.
    \subitem \nameref{cmd:reg-read}: \texttt{0x11 + 0xED\_12\_3\hl{X}\_0xF8}
  \item Verify the memory contents of N3.
  \item N1 requests that 9 of the registers are written to N2:
    \subitem \nameref{cmd:mem-stream-read}: \texttt{\hl{0x??}}
\end{enumerate}

\subsection{Memory}

\subsubsection{Bulk Transaction Overflow Wrapping}
\label{test:mem-bulk-overflow}

\begin{quote}
  \texttt{A) 3 Nodes: N1 -> N2 -> MED (-> N1)}
\end{quote}

This test verifies both basic memory bulk transfer operation and that the bulk
transfers handle overflow wrapping correctly.

\begin{enumerate}
  \item Begin with N1's memory full of known, non-zero values. N2 memory
    should be zeroed.
  \item N1 sends a request to N2 to read the last word and first 3 words
    of its memory and writes it to the last 2 words and first 2 words of N1's
    memory.
    \subitem \nameref{cmd:mem-bulk-read}: \texttt{0x25 + 0xfffffffc\_12\_000011\_ffffffff8}
  \item Verify the correct values from N2 are written to memory in N1.
\end{enumerate}

\begin{verbatim}
  Test Name:            Success:
  mem-bulk-overflow-A - Write 4 words to N2
\end{verbatim}

\subsection{Interjection Timing}

\subsubsection{Third-Party Interjector for Many Registers}
\label{test:inj-reg-long}

\begin{quote}
  \texttt{A) 4 Nodes: RX -> INJ -> TX -> MED (-> RX)}
\end{quote}

Have the TX node send a register write command that writes at least
3~registers long. The INJ node should generate an interjection on the \nth{63}
bit of {\bf data} with control bit~0 set to {\tt !EoM}. The effect should be
that the transmitter sees 63 data clock edges and the receiver sees 65 data
clock edges. At the end of the transaction, the first register should be
written, the second register should not. The TX layer controller should
believe that the transmission failed and it has written either 7~bytes or
1~word (depending on implementation), but not 8~bytes or 2~words.

Repeat this test, interjecting on the 64th edge. The result should be the
same.

\begin{quote}
  \texttt{B) 4 Nodes: TX -> INJ -> RX -> MED (-> TX)}\\
  \texttt{C) 4 Nodes: TX -> RX -> INJ -> MED (-> TX)}
\end{quote}

Repeat the same test with these node positions.

Repeat these tests, interjecting on the 64th edge. The result should be the
same.

\begin{quote}
  \texttt{D) 4 Nodes: INJ -> RX -> TX -> MED (-> INJ)}
\end{quote}

Repeat the same test but interject at the \nth{64}, \nth{65}, and \nth{66}
data edges. For 64 and 65, the result should be the same. For 66, two
registers should be written.

\begin{verbatim}
  Test Name:          Success:
  inj-reg-long-A-63 - Write 1 register
  inj-reg-long-A-64 - Write 1 register
  inj-reg-long-B-63 - Write 1 register
  inj-reg-long-B-64 - Write 1 register
  inj-reg-long-C-63 - Write 1 register
  inj-reg-long-C-64 - Write 1 register
  inj-reg-long-D-63 - Write 1 register
  inj-reg-long-D-64 - Write 1 register
  inj-reg-long-D-65 - Write 1 register
  inj-reg-long-D-66 - Write 2 registers
\end{verbatim}

\subsubsection{Third-Party Interjector at Last Register}
\label{test:inj-reg-end}

This test is the same as the previous, except the TX node should attempt to
write exactly 2~registers.

\begin{verbatim}
  Test Name:         Success:
  inj-reg-end-A-63 - Write 1 register
  inj-reg-end-A-64 - Write 1 register
  inj-reg-end-B-63 - Write 1 register
  inj-reg-end-B-64 - Write 2 registers
  inj-reg-end-C-63 - Write 1 register
  inj-reg-end-C-64 - Write 2 registers
  inj-reg-end-D-63 - Write 1 register
  inj-reg-end-D-64 - Write 1 register
  inj-reg-end-D-65 - Write 2 registers
  inj-reg-end-D-66 - Write 2 registers
\end{verbatim}

\subsubsection{Third-Party Interjector for Long Memory Bulk Transfer}
\label{test:inj-mem-bulk-long}

Repeat the same tests as \nameref{test:inj-reg-long}, adding an address to the
\bus data payload and interjecting 32~cycles later such that either one or two
words are written to memory.

\begin{verbatim}
  Test Name:          Success:
  inj-mem-bulk-long-A-63 - Write 1 word
  inj-mem-bulk-long-A-64 - Write 1 word
  inj-mem-bulk-long-B-63 - Write 1 word
  inj-mem-bulk-long-B-64 - Write 1 word
  inj-mem-bulk-long-C-63 - Write 1 word
  inj-mem-bulk-long-C-64 - Write 1 word
  inj-mem-bulk-long-D-63 - Write 1 word
  inj-mem-bulk-long-D-64 - Write 1 word
  inj-mem-bulk-long-D-65 - Write 1 word
  inj-mem-bulk-long-D-66 - Write 2 words
\end{verbatim}

\subsubsection{Third-Party Interjector at End of Memory Bulk Transfer}
\label{test:inj-mem-bulk-end}

Repeat the same tests as \nameref{test:inj-reg-end}, adding an address to the
\bus data payload and interjecting 32~cycles later such that either one or two
words are written to memory.

\begin{verbatim}
  Test Name:         Success:
  inj-mem-bulk-end-A-63 - Write 1 word
  inj-mem-bulk-end-A-64 - Write 1 word
  inj-mem-bulk-end-B-63 - Write 1 word
  inj-mem-bulk-end-B-64 - Write 2 words
  inj-mem-bulk-end-C-63 - Write 1 word
  inj-mem-bulk-end-C-64 - Write 2 words
  inj-mem-bulk-end-D-63 - Write 1 word
  inj-mem-bulk-end-D-64 - Write 1 word
  inj-mem-bulk-end-D-65 - Write 2 words
  inj-mem-bulk-end-D-66 - Write 2 words
\end{verbatim}
